Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down

### 2022-06-08 paper/ms.tex
Other small changes

A clear prediction of this proposed mechanism is that the constant pattern
speed a particular galaxy will end up is somewhat arbitrary. In the real
universe for an isolated galaxy it would be the formation pattern speed of the
bar while in our simulation it is the pattern speed of the bar when gas is
added to the system. We tested this by adding gas to the system at a later
time when the bar has further grown and slowed down with time. In our
particular test, we added the gas at a time when the pattern speed is
$\sim30\,\textrm{km}/\textrm{s}/\textrm{kpc}$. As shown in
Fig.~\ref{fig:snap700}, we find that the pattern speed evolution is very
similar between the two cases (orange and red lines). When the initial pattern
speed is lower (red line), the addition of gas does not cause the pattern
speed to quickly return to the higher value of our fiducial simulation (orange
line).

@@

We computed the circular velocity curve of our model using the \texttt{AGAMA}
package.\cite{2019MNRAS.482.1525V} We fit the baryonic component (stellar
disk, bulge, gas, and newly formed stars) with an axisymmetric cylindrical
spline with $20$ grid points in both the radial and vertical direction
spanning $0.2$ to $50\,\textrm{kpc}$ in the radial direction and from $0.02$
to $10\,\textrm{kpc}$ in the vertical direction. We fit the dark matter halo
using a spherically symmetric multipole fit with a maximum angular harmonic
coefficient of $l=2$. We plot the circular velocity curve in Extended Data
Fig.~\ref{fig:vcirc} compared to observational
estimates.\cite{2019ApJ...871..120E} Our disk is slightly more massive than
the Milky Way disk, and in the SMUGGLE run the addition of the gas phase
results in a slightly higher circular velocity. We also show the evolution of
the surface density profile in Extended Data Fig.~\ref{fig:surf} We find that
in our simulation the atomic and molecular gas surface density and the SFR
surface density is broadly consistent with the expected values for the Milky
Way.\cite{2008AA...487..951K,2022ApJ...929L..18E} The discrepancy between $1$
and $4\,\textrm{kpc}$ in the molecular and SFR surface density is likely due
to the fact that the distances to molecular clouds which underlines this work
used a simple kinematic distance based on an axisymmetric model of the Milky
Way,\cite{2017ApJ...834...57M} which is not accurate in the bar region
where gas has large non-circular velocities.

>>>>

A clear prediction of this proposed mechanism is that the constant pattern
speed a particular galaxy will end up is somewhat arbitrary. In the real
universe for an isolated galaxy it would be the formation pattern speed of the
bar while in our simulation it is the pattern speed of the bar when gas is
added to the system. We tested this by adding gas to the system at a later
time when the bar has further grown and slowed down with time. In our
particular test, we added the gas at a time when the pattern speed is
$\sim30\,\textrm{km}/\textrm{s}/\textrm{kpc}$. As shown in
Fig.~\ref{fig:snap700}, we find that the pattern speed evolution is very
similar between the two cases (orange and red lines). If anything, the system
with a lower pattern speed seems to speed up more, which is consistent with
our picture since the stronger bar should experience a larger torque from the
gas as it is more efficient at driving gas inflows. We also show in the
Methods that more slowly rotating bars at fixed bar strength are more
efficient at driving as inflows as well. Nonetheless, when the initial pattern
speed is lower (red line), the addition of gas does not cause the pattern
speed to quickly return to the higher value of our fiducial simulation (orange
line).

@@

We computed the circular velocity curve of our model using the \texttt{AGAMA}
package.\cite{2019MNRAS.482.1525V} We fit the baryonic component (stellar
disk, bulge, gas, and newly formed stars) with an axisymmetric cylindrical
spline with $20$ grid points in both the radial and vertical direction
spanning $0.2$ to $50\,\textrm{kpc}$ in the radial direction and from $0.02$
to $10\,\textrm{kpc}$ in the vertical direction. We fit the dark matter halo
using a spherically symmetric multipole fit with a maximum angular harmonic
coefficient of $l=2$. We plot the circular velocity curve in Extended Data
Fig.~\ref{fig:vcirc} compared to observational
estimates.\cite{2019ApJ...871..120E} Our disk is slightly more massive than
the Milky Way disk, and in the SMUGGLE run the addition of the gas phase
results in a slightly higher circular velocity.

We also show the evolution of the surface density profile in Extended Data
Fig.~\ref{fig:surf} We find that in our simulation the atomic and molecular
gas surface density and the SFR surface density is broadly consistent with the
expected values for the Milky
Way.\cite{2008AA...487..951K,2022ApJ...929L..18E} The discrepancy between $1$
and $4\,\textrm{kpc}$ in the molecular and SFR surface density is likely due
to the fact that the distances to molecular clouds which underlines this work
used a simple kinematic distance based on an axisymmetric model of the Milky
Way,\cite{2017ApJ...834...57M} which is not accurate in the bar region where
gas has large non-circular velocities.
